Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-03-03 index.qmd
updated methods

@@

$$
h(t|\mathbf{Z}) = h_0(t)\text{exp}(\Sigma_{k=1}^{p} \beta_kZ_k)
$$ {#eq-hazard}

-   $h(t|\mathbf{Z})$ is the hazard rate at time $t$ for an individual
    with risk vector $\mathbf{Z}$
-   $h_0(t)$ is an arbitrary baseline hazard rate
-   $\mathbf{Z}$ is the set of $p$ covariates
-   $\beta$ is the set of $p$ coefficients that measure the impact of
    the covariates

@@

The proportional hazards assumption has been proven to be critical to evaluate prior to using Cox Proportional Hazards when modeling survival data. The proportional hazards assumption states the ratio of hazard rates for any two individuals is constant over time regardless of the point in time of the measurement meaning the effect of the covariates must be constant over time. The assumption will be tested by calculating scaled Shoenfeld residuals and plotting them against time. The resulting plot for each covariate is interpreted to prove the assumption is met if the plot has Shoenfeld residuals roughly constant over time. [@Bustan2018]

This assumption describes how if an event is censored it is unrelated to the model being created. Censored data is difficult to accommodate when Cox proportional hazards models are used in other industries, such as the medical field, because observation times for the individuals involved in studies may have decades before an event is observed. In this scenario, it would not be appropriate to consider this individual when modeling survival. In the dataset being used for this model, the training data does not contain censored data. The documentation available at the dataset's source clarifies all engines in the training data have experienced the desired event of failure.

@@

**Checking Proportional Hazards**
```{r}
cox.zph(cox)

@@

>>>>

CPH relies on data that has a time to event component and covariates for
that event.

The proportional hazards assumption has been proven to be critical to
evaluate prior to using CPH when modeling survival data. The
proportional hazards assumption states that the ratio of hazard rates
for any two individuals must be constant over time regardless of the
point in time of the measurement [@Bustan2018] . This means the effect
of the covariates must be constant over time. This assumption can be
tested using the R package `survival`. The package contains a
proportional hazards test called `cox.zph()` [@R-base]. Using the
hypothesis,

$H_0$: the predictor's coefficient does not vary with time

$H_1$: the predictor's coefficient does vary with time

the `cox.zph()` calculates the scaled Schoenfeld partial residuals and
returns the $\chi^2$ value, the degrees of freedom, and p-values for
each individual covariate along with a global value. Using an
$\alpha = 0.05$ and the global p-value, the determination of whether to
reject or fail to reject $H_0$ can be made [@nahhas2025]. Individual
plots for each covariate can then be generated which provide a visual
interpretation of the p-values generated from `cox.zph()`.

This assumption describes how if an event is censored it is unrelated to
the model being created. Censored data is difficult to accommodate when
Cox proportional hazards models are used in other industries, such as
the medical field, because observation times for the individuals
involved in studies may have decades before an event is observed. In
this scenario, it would not be appropriate to consider this individual
when modeling survival.

### Equations

$$
h(t|\mathbf{Z}) = h_0(t)\text{exp}(\sum_{k=1}^{p} \beta_kZ_k)
$$ {#eq-hazard}

-   $h(t|\mathbf{Z})$ is the hazard rate at time $t$ for an individual
    with risk vector $\mathbf{Z}$
-   $h_0(t)$ is an arbitrary baseline hazard rate
-   $\mathbf{Z}$ is the set of $p$ covariates
-   $\beta$ is the set of $p$ coefficients that measure the impact of
    the covariates [@klein2005]

CPH is a proportional hazards model because the ratio of hazard rates
between two individuals is

$$
\frac{h(t|\mathbf{Z})}{h(t|\mathbf{Z}^*)} = \frac{h_0(t)\text{exp}(\Sigma_{k=1}^{p} \beta_kZ_k)}{h_0(t)\text{exp}(\Sigma_{k=1}^{p} \beta_kZ_k^*)} = \text{exp}[\sum_{k=1}^p \beta_k(Z_k-Z_k^*)]
$$ {#eq-hazardsratio}

-   $\mathbf{Z}$ and $\mathbf{Z}^*$ are covariate values [@klein2005]

@@

@@

@@

**Checking Proportional Hazards**

```{r}
cox.zph(cox)

@@

**Checking Independent Censoring**

In the dataset being used for this model, the training data does not
contain censored data. The documentation available at the dataset's
source clarifies all engines in the training data have experienced the
desired event of failure.
